# Human Activity Recognition on Smartphones using a Tree Approach

```{r libraries}

library(ggplot2)
```


```{r loaddata}

setwd("C:/Users/alem/Documents/Data Analysis/Samsung Phone Data Analysis")
load("./data/samsungData.rda")
```

## Exploratory analysis

### Check for NAs


```{r expanalysisNAs}

dim(samsungData)
sum(complete.cases(samsungData))
```

* No NAs included

### Check for required transformations

```{r expanalysisTrans, results='hide'}

names(samsungData)
str(samsungData, list.len = 563)

samsungData$subject <- as.factor(samsungData$subject)
samsungData$activity <- as.factor(samsungData$activity)

```

* The *subject* and the *activity* variable were converted into factors

### Exploring the subject variable

```{r expanalysisSubject}

length(unique(samsungData$subject))
unique(samsungData$subject)
table(samsungData$subject)
ggplot(samsungData, aes(subject)) + geom_bar()
mean(table(samsungData$subject))
```

* In total, data from 21 different subjects is included in the data set
* The 21 different subjects were labeled as follows: `r unique(samsungData$subject)`
* On average `r mean(table(samsungData$subject))` observations were collected for each subject
* The number of observations per subject is nearly uniformly distributed

### Exploring the acitivity variable

```{r expanalysisActivity}

length(unique(samsungData$activity))
unique(samsungData$activity)
table(samsungData$activity)
mean(table(samsungData$activity))
ggplot(samsungData, aes(activity)) + geom_bar()
```

* In total, data from 6 different activities is included in the data set
* The 6 activities are the following: `r unique(samsungData$activity)`
* On average `r mean(table(samsungData$activity))` observations were collected for each activity
* For activity *walkdown* and *walkup* slightly fewer observations are available in the data set when compared to the average (`r mean(table(samsungData$activity)) - as.numeric(table(samsungData$activity)[5])` and `r mean(table(samsungData$activity)) - as.numeric(table(samsungData$activity)[6])` respectively)

## Creating training set, validation test, and test set

The traning set will be used for training the models, the validation set for picking models, and the test set for reporting the final error value.

Training set includes:
* (obs: 1, 3, 5, 6, 7, 8, 11, 14)

Validation set includes:
* (obs: 15, 16, 17, 19, 21, 22, 23)

Test set includes:
* (obs: 27, 28, 29, 30)

```{r subset}

training_set <- subset(samsungData, subject %in% c(1, 3, 5, 6, 7, 8, 11, 14))
unique(training_set$subject)

validition_set <- subset(samsungData, subject %in% c(15, 16, 17, 19, 21, 22, 23))
unique(validition_set$subject)

test_set <- subset(samsungData, subject %in% c(27, 28, 29, 30))
unique(test_set$subject)

```







